---
title: "Reverse Engineering Project"
author: "Dylan, Oliver, Michael"
date: "10/06/2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Many American's Don't Vote ](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(tidyverse)
library(janitor)
library(dplyr)
library(refinr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
american_voter_data<- read_csv("data/nonvoters_data.csv")
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn't matter or isn't worth their time.
**Analysis summary**: We connected Q26 to Q29 to confirm the finding that either their vote does not matter or that voting is not worth their time. We were able to confirm this based on the responses to Q26 with prompts 2,3, and 4 in the nonvoters codebook PDF.

```{r}

filtered_data <- american_voter_data|>
filter (Q26 >= 2) |>
filter (Q29_2 == 1)
print(filtered_data)


```

### Sentence 2

* **Sentence text**: In the survey, we asked voters who have missed at least one national election - which included some people who almost always vote - why they didn't cast a ballot. Nearly a quarter cited some of the structural barriers we mentioned above. But another 31 percent said that they decided not to vote because they disliked the candidates or they thought nothing would change as a result of the election (26 percent).
* **Analysis summary**: Out of 1,342 people who say they sometimes, rarely, or don't vote in national elections, 765 say that they either dislike the candidates or that nothing will change for them. 403 say they dislike the candidates, while 362 say nothing will change as a result. This is
similar to the original numbers presented in the text, as 403 is 30% of 1342 and 362 is 27%.

```{r}
filtered_data <- american_voter_data|>
  filter(Q26==2 | Q26==3|Q26==4)
                    
filtered_data
filtered_data<-filtered_data|>
  filter(Q29_1 == 1 | Q29_3 == 1)
filtered_data
```

### Sentence 3

* **Sentence text**: They were also likelier than those who vote more regularly to say they couldn’t get off work to vote...
Also because he doesn’t trust mail-in ballots. “The last time I tried to vote by mail, it didn’t get counted — it was too late,” he said. As a result, Sanchez told us, “I don't know where my ballot is going to end up, if I don't see it put into the voting machine.”
* **Analysis summary**: When filtering the data, 229 of the respondents chose 5 on the survey for Q29, which has the answer as they wanted to vote but had work/ other obligations or forgot. Furthermore, question 19 asks respondents what they think would make it easier for them to vote. I filtered the responses that have to do with automatically being registered to vote and automatically receiving a mail- in ballot which had thousands of respondents say this would make voting a lot more appealing for them. Comparing these different answers shows what would make voting more easier for those who might not know about mail-in deadlines.

```{r}

race_data <- american_voter_data|>
  select(Q29_5)|>
filter(Q29_5 == 1)
race_data
```


```{r}
race_data <- american_voter_data|>
  select(Q19_4)|>
filter(Q19_4 == 1)
race_data
```

```{r}
race_data <- american_voter_data|>
  select(Q19_5)|>
filter(Q19_5 == 1)
race_data
```


### Sentence 4

* **Sentence text**: A substantial chunk of respondents said they don’t think either party wants people like them to vote. (Twenty-three percent said this of Democrats and 31 percent of Republicans.) Black and Hispanic people were especially likely to say the Republicans don’t want people like them to vote: Fifty-four percent of Black respondents and 35 percent of Hispanic respondents agreed with this statement, compared to just 26 percent of white respondents.

* **Analysis summary**: While this would be the correct way to find the the percentages of people who think each party does/ does not care about their vote, I think one of the issues is that this data set does not specify which race answered to each question. The republican party shows that 33% of respondents say they don't think republicans want them to vote, similar to the 31% in the article. Q15, which is the democratic party, got a number of 21%, but accounting for the people that didn't vote the number should be similar to the 23% in the article.

```{r}
american_voter_data|>
group_by(Q14)|>
summarise (count=n())|>
mutate(percent=count/sum(count)*100)
```

```{r}
american_voter_data|>
group_by(Q15)|>
summarise (count=n())|>
mutate(percent=count/sum(count)*100)
```

### Sentence 5

* **Sentence text**:Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.”
* **Analysis summary**: The first part shows how many Black people said they had to stand in line for more than an hour and the same for Hispanic voters in the next code section. Then you just delete the filter q18 part to get the total amount of black people. For each output, there are over 100 respondants who have said they waited in line for over an hour to vote. 

```{r}
filtered_data<- american_voter_data
  filtered_data|>
  filter(race== 'Black')|>
  filter(Q18_8 == 1)|>
  print(filtered_data)

  filtered_data<- american_voter_data
  filtered_data|>
    filter(race == "Hispanic") |>
    filter(Q18_2== 1 | Q18_7 ==1)|>
    print(filtered_data)
```

-30-