---
title: "Reverse Engineering Project"
author: "Dylan, Oliver, Michael"
date: "10/06/2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Many American's Don't Vote ](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(tidyverse)
library(janitor)
library(dplyr)
library(refinr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
american_voter_data<- read_csv("data/nonvoters_data.csv")
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn't matter or isn't worth their time.
**Analysis summary**: We connected Q26 to Q29 to confirm the finding that either their vote does not matter or that voting is not worth their time. We were able to confirm this based on the responses to Q26 with prompts 2,3, and 4 in the nonvoters codebook PDF.

```{r}
# Put code to reverse engineer sentence here
american_voter_data|>
  filter(Q26)|>
  group_by("2,3,4")|>
  summarise(Q26)


# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-