---
title: "lab_01"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r}
winred_wv_data <- read_rds("lab_01.rds") 
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.

**Q1** How many rows and columns does this dataframe have?

**A1** 40,441 rows 24 columns

```{r}
glimpse(winred_wv_data)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?

**A2**Charleston, 3226

```{r}
winred_wv_data |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))
```

**Q3** Which West Virginia cities accounted for the fewest contributions? How many of them, roughly speaking?

**A3**63 cities

```{r}
winred_wv_data |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange((count_contribs))
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4** I looked up the populations in the top 5 of the cities listed and all of them have very low populations, with some only having around 30 people. This most likely means that there are simply not enough people in the town to make any contributions 

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5** Trump National Committee- they had 14,063 contributions and the average donation was 32.99

```{r}
winred_wv_data |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)
  ) |>
  arrange(desc(count_contribs))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6** The Trump National Committee was the top recipient. This makes sense since West Virginia is a majority Republican state. According to 270towin.com, for the 2020 presidential election the results ended up being 68.6% of voters voting Republican 

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**.

```{r}

wv_summary<-winred_wv_data |>
group_by(memo_text) |>
  summarise( 
    count_contributions = n(),
    total_amount = sum(amount),
    avg_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)) 
```
